Detecting Facial Expressions in Professional Tennis Matches

### 2016-12-16 Report06-12.Rmd
Refining Analysis script

---
title: "Detecting Facial Expressions in Professional Tennis Matches"
author: Stephanie Kobakian, Mitchell O'Hara Wild. Supervised by Dianne Cook, Stephanie
  Kovalcik
date: "November 19 2016"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
    fig_caption: yes
  word_document:
    toc: yes
fontsize: 11pt
bibliography: references.bib
---

@@

knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE,
                      cache=TRUE, fig.align = 'center')

@@

#plot images with overlaid boxes
overlayGgplot <- function(imgList, mergeData, matchBox=TRUE){
  for(i in imgList){

@@

    faceData <- ALLmetaIMG %>% filter(file == i)

@@

    }
    if(matchBox){
      p <- p + geom_path(aes(x=x, y=450-y, group = faceID, colour=faceID), faceData) + guides(colour="none")
    }
    else{
      p <- p + geom_path(aes(x=x, y=450-y, group = faceID, colour=type), faceData)
 }
  }
  return(p)
}

@@

@@

We were able to try the online demos to see whether these softwares were viable, we used the following image:
```{r Trial Image, results=FALSE}
imagesList0<-as.list("2016_HSA_R01_BTomic_AUS_vs_DIstomin_UZB_MS157_clip.0013.png")
overlayGgplot(imagesList0, ALLmetaIMG, matchBox = FALSE)
```

## Methodology ##
An annotation tool was constructed to create a base for comparative analysis, we refer to this as Manual Classifications.
These manual Classifications involved describing the features of the Scenes. For this we considered the following attributes of the scenes:

#####**Table 1.** Scene Attributes
```{r Scene-Attributes-Table, anchor="table"}
kable(SceneAttributes, format = "pandoc", booktabs = TRUE, caption = figr("Scene-Attributes-Table", TRUE, type="Table"))
```

`r figr("Scene-Attributes-Table", TRUE, type="Table")`   Yayyy

The aim was to collect specific information on each different face within the scene. To determine which of the sometimes many faces in the scene it would be reasonable for softwares to detect a standard was created for reasonable detection. This was based on the attributes provided in the table below:

#####**Table 2.** Solution Specifications
```{r Solution-Specifications-Table, results='asis', anchor="table"}
kable(SolutionSpecs, format = "pandoc", booktabs = TRUE)

@@

#####**Table \label{tab:FAT}.** Face Attributes
```{r T3 Face Attribute Tables, results='asis'}
kable(FaceAttributes, format = "pandoc", booktabs = TRUE)
```

Table \ref{tab:FAT} above

####Manual Annotation
To record these questions a Shiny App was created. We called this Application our ManualClassificationProgram. This helped to provide information for all attributes readily and consistently. This worked by presenting an image, using the imager package, from the sample of images that had not yet been considered, appearing underneath were a set of radio buttons corresponding to the Scene table as shown above. If there was a face in the image we were able to highlight the square 'Face Box'. This recorded the x and y coordinates of the box drawn, and when the save button was hit it saved all the radio button selections and the 'Face Box' coordinates to a CSV file.

@@

#####**Graph 1.** Face Bounding Box Data: Bar Chart
```{r G1 Face Bounding Box Data, results='hide'}
ggplot(ALLmetaIMG, aes(x = type, fill=type)) + geom_bar(position="dodge")  +xlab("Facial Recognition Softwares")+scale_x_discrete(limits=c("Manual","Google","Microsoft","Skybiometry","Animetrics")) +ylab("Number of Potential Faces") + guides(fill=FALSE) 
```

@@

#####**Graph 4.** Face Bounding Box Data: Box Plot 
```{r G4 Face Bounding Box Data}
ggplot(ALLmetaIMG, aes(x = factor(x = factor(type)), fill=type, y = size)) + geom_boxplot() + scale_y_continuous(trans='log') +xlab("Facial Recognition Softwares")+scale_x_discrete(limits=c("Google","Animetrics","Manual","Microsoft","Skybiometry"))+ylab("Size of Face Bounding Boxes")+guides(fill=FALSE)

@@

##### Image 2.2016_CT6_R02_JSousa_POR_vs_SGiraldo_COL_MS231_clip.0017.png
```{r I2, results=TRUE}
imagesList2<-as.list("2016_CT6_R02_JSousa_POR_vs_SGiraldo_COL_MS231_clip.0017.png")
overlayGgplot(imagesList2, ALLmetaIMG, matchBox = FALSE)
```

@@

##### Image 3.2016_SC2_R01_ATomljanovic_AUS_vs_KBondarenko_UKR_WS1112_clip.0053.png
```{r I3, results=TRUE}
imagesList3<-as.list("2016_SC2_R01_ATomljanovic_AUS_vs_KBondarenko_UKR_WS1112_clip.0053.png") 
overlayGgplot(imagesList3, ALLmetaIMG, matchBox = FALSE)
```

@@

##### Image 4.2016_RLA_R04_NDjokovic_SRB_vs_GSimon_FRA_MS401_clip.0003.png
```{r I4,results=TRUE}
imagesList4<-as.list("2016_RLA_R04_NDjokovic_SRB_vs_GSimon_FRA_MS401_clip.0003.png")
overlayGgplot(imagesList4, ALLmetaIMG, matchBox = FALSE)
```

This is almost usable, however this shows the sensitivity of the software. We may consider this too sensitive as it recognised the eye and nose of the ballboy on the court.

##### Image 5.2016_SC2_R01_AKerber_GER_vs_MDoi_JPN_WS1332_clip.0083.png
```{r I5, results=TRUE}
imagesList5<-("2016_SC2_R01_AKerber_GER_vs_MDoi_JPN_WS1332_clip.0083.png") 
overlayGgplot(imagesList5, ALLmetaIMG, matchBox = FALSE)
```

@@

#####**Table 4.** Type Matches Manual
```{r T4 Type Matches Manual}
library(descr, quietly = TRUE)
NoManual<-ALLmetaIMG %>% filter(ALLmetaIMG$type!="Manual")
NoManual<-NoManual %>% droplevels(NoManual$type)
Accuracy<-with(NoManual, table(NoManual$matchesManual, NoManual$type))
pander(CrossTable(Accuracy, prop.chisq = FALSE) , digits = 2)
```

@@

#####**Table 5.** Image Characteristics
```{r T5 Image Characteristics Table}
ImageCharacteristics <- ALLmetaIMG %>% filter(type=="Manual") %>% group_by(person, situation, bg, shotangle, detect) %>% dplyr::summarise(count=n())
ImageCharacteristics<-arrange(ImageCharacteristics, desc(count))

kable(ImageCharacteristics[1:10,], format = "pandoc", booktabs = TRUE)
```

This table displays the feature combinations that produced the most potential face Bounding Boxes recognitions by all four softwares. The most common shot is a crowd shot.
The second row in the table with 830 faces recognised is more interesting than the first result. This useful scene is an image of a Court Player Close-Up in front of a Logo Wall, taken at Player Shoulder Height.

@@

##### **Image 1.Typically Recognised Image**

```{r I1 Typical Best Recognised, results=FALSE}
imagesList1<-as.list("2016_CT6_R01_JMillman_AUS_vs_DSchwartzman_ARG_MS159_clip.0013.png") 
overlayGgplot(imagesList1, ALLmetaIMG, matchBox = FALSE)
```

#####**Table 6.** Google Errors Characteristics
```{r T6 Google Errors Characteristics Table}
library(dplyr)

@@

GoogleNoMatch<-arrange(GoogleNoMatch, desc(count))
pander(GoogleNoMatch[1:10,], type = 'table')

@@

This Table shows that the potential Face Bounding Boxes Google returned were found in images that had the same characteristics of those that were manually annotated for faces.
This is shown by the same combinations of characteristics occuring in both Table 5 and Table 6. 
Without looking at each individual image this Table confirms that the potential Face Bounding Boxes Google located will likely be reasonable, and actually contain faces.

The best scenes for facial recongition have been found, given this information the following table considers the Characteristics of the individual faces found within those scenes. For this section we chose to consider only the faces that were manually annotated as players, with the intention of not basing results on recognitions of undesirable faces.

#####**Table 7.** Face Characteristics
```{r T7 Face Characteristics Table}
FaceCharacteristics <- ALLmetaIMGPlayers %>%   filter(type=="Google")%>% filter(matchesManual==TRUE)%>% group_by(visorhat, glasses, headangle, lighting, obscured, detect) %>% dplyr::summarise(count=n())
FaceCharacteristics<-arrange(FaceCharacteristics, desc(count))
pander(FaceCharacteristics[1:10,], type = 'table')
```

Table 7 utilises a set of faces that were Manually annotated and also found by Google. 
Nine of the top ten facial characteristic combinations contained faces that were not wearing glasses.
The head angle describing the face angle was 'Other'^[Definition for headangle factor level 'Other' found in..]
for nine of the top ten facial characteristic combinations.

#####**Table 8.** Glasses on Recognised Faces
```{r T8 Glasses}
GlassesGoogle <- ALLmetaIMGPlayers %>% filter(type=="Google") %>%
filter(matchesManual==TRUE)

@@

#####**Graph 2.** Glasses
```{r G2 Glasses}
ggplotProportion(ALLmetaIMGPlayers, glasses)
```
Graph 2 above shows that Google outperforms the other softwares, with or without glasses. When there are no glasses worn Google finds over 60% of the manually annotated faces.

#####**Table 9.** Head Angles
```{r T9 HeadAngles}
HeadangleGoogle <- ALLmetaIMGPlayers %>%   filter(type=="Google") %>%
filter(matchesManual==TRUE)

pander(table(HeadangleGoogle$headangle))
```
The amount of faces found by Google with the headangle "Other", is much larger than the amount of faces with the headangle "Profile" or "Front On". Therefor this should also be considered propotionately.

#####**Graph 3.** Glasses
```{r G3 Glasses}
ggplotProportion(ALLmetaIMGPlayers, headangle)
```
Graph 3 diplays that Google performs much better in comparison to the other softwares when the headangle is Profile. This is outperforming unusually well, however this could also be due to the poor performance of the softwares in this circumstance.

@@

##### **Table 10.**
```{r T10}
library(tables, quietly = TRUE)
Accuracy<-with(ALLmetaIMG, table(ALLmetaIMG$matchesManual, ALLmetaIMG$type))
pander(Accuracy)
```

@@

#####**Table 11.** Google Characteristics
```{r T11 Google Characteristics Table}
GoogleCharacteristics <- ALLmetaIMG %>% filter(type=="Google") %>% filter(person=="Person") %>% filter(situation=="Court player close-up") %>% filter(bg=="Logo wall") %>% filter(detect=="Player") %>% group_by(person, situation, bg, shotangle, detect) %>% dplyr::summarise(count=n())
GoogleCharacteristics<-arrange(GoogleCharacteristics, desc(count))

kable(GoogleCharacteristics, format = "pandoc", booktabs = TRUE)
```
This table shows how many images with potential player's faces Google recognised. This diplays a vast gap between the amount of potential faces found given the different shot angles. Shoulder height is an optimum angle for a Face Bounding Box.

@@

#####**Graph 4.** Google Face Bounding Boxes: Mosaic Plot
```{r G4 Google Characteristics Mosaic Plot}
GoogleAccessories <- ALLmetaIMG %>% filter(type=="Google") %>% filter(shotangle=="Player Shoulder Height") %>% filter(detect=="Player") %>% filter(lighting!="Direct sunlight")

GoogleAccessories<-droplevels(GoogleAccessories)
#lighting headangle glasses visorhat

mosaicplot(~ headangle + visorhat, data = GoogleAccessories, color = TRUE)

mosaicplot(~ lighting + visorhat, data = GoogleAccessories, color = TRUE)

mosaicplot(~ lighting +headangle +visorhat, data = GoogleAccessories, color = TRUE)

mosaicplot(~ glasses + lighting, data = GoogleAccessories, color = TRUE)

>>>>

---
title: "Detecting Facial Expressions in Professional Tennis Matches"
author: Stephanie Kobakian, Mitchell O'Hara-Wild. Supervised by Dianne Cook, Stephanie
  Kovalcik
date: "November 19 2016"
output:
  pdf_document:
    fig_caption: yes
    toc: yes
    toc_depth: 2
fontsize: 10pt
bibliography: references.bib
---

@@

knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, kfigr.link=TRUE)

@@

#plot images with overlaid boxes
overlayGgplot <- function(imgList, mergeData, matchBox=FALSE){
  for(i in imgList){

@@

    faceData <- ALLmetaIMGFaces %>% filter(file == i)

@@

    }
    if(matchBox){
      p <- p + geom_path(aes(x=x, y=450-y, group = type, colour=boxID), faceData) + guides(colour="none")
    }
    else{
      p <- p + geom_path(aes(x=x, y=450-y, group = type, colour=type), faceData)
    }
  return(p)
  }
}

@@

#change this to facesVenn with binary information
ALLmetaIMGFaces<-ALLmetaIMG%>%
  filter(!is.na(boxID)) %>% 
  select(-ID, -facecounter)%>%
  mutate(fileID = as.numeric(factor(file))) %>%
  mutate(FaceKey=paste(fileID, boxID, sep="-")) %>%
  mutate(FaceID=paste(fileID, boxID, substring(type, 1,2), sep="-"))

@@

We were able to try the online demos to see whether these softwares were viable, we used the following  `r figr("Trial-Image", TRUE, type="Figure")`:

```{r Trial-Image, fig.cap="This is a player", anchor="figure"}
imagesList0<-as.list("2016_HSA_R01_BTomic_AUS_vs_DIstomin_UZB_MS157_clip.0013.png")
overlayGgplot(imagesList0, ALLmetaIMG, matchBox = FALSE)
```

## Methodology ##
An annotation tool was constructed to create a base for comparative analysis, we refer to this as Manual Classifications.
These manual Classifications involved describing the features of the Scenes.

```{r Scene-Attributes-Table, anchor="table", fig.cap="Attributes of a Scene of an image. The most appropriate option from the list was selected for each attribute."}
kable(SceneAttributes, format = "pandoc", booktabs = TRUE)
```

The aim was to collect specific information on each different face within the scene. To determine which of the sometimes many faces in the scene it would be reasonable for softwares to detect a standard was created for reasonable detection. This was based on the attributes provided in `r figr("Solution-Specifications-Table", TRUE, type="Table")`, below:

```{r Solution-Specifications-Table, results='asis', anchor="table"}
kable(SolutionSpecs, format = "pandoc", booktabs = TRUE, caption = "Software solutions specifications.")

@@

```{r Face-Attribute-Tables, results='asis', anchor="table"}
kable(FaceAttributes, format = "pandoc", booktabs = TRUE, caption = "Attributes of an individual face within an image. The most appropriate option from the list was selected for each attribute.")
```

####Manual Annotation
To record these questions a Shiny App was created. We called this Application our ManualClassificationProgram. This helped to provide information for all attributes readily and consistently. This worked by presenting an image, using the imager package, from the sample of images that had not yet been considered, appearing underneath were a set of radio buttons corresponding to the Scene Attibutes in `r figr("Scene-Attributes-Table", TRUE, type="Table")` as shown above. If there was a face in the image we were able to highlight the square 'Face Box'. This recorded the x and y coordinates of the box drawn, and when the save button was hit it saved all the radio button selections and the 'Face Box' coordinates to a CSV file.

@@

```{r Face-Bounding-Box-Data, results='hide', type="figure", fig.cap='..', anchor="figure"}
ggplot(ALLmetaIMG, aes(x = type, fill=type)) + geom_bar(position="dodge") +xlab("Facial Recognition Softwares")+scale_x_discrete(limits=c("Manual","Google","Microsoft","Skybiometry","Animetrics")) +ylab("Number of Potential Faces") + guides(fill=FALSE)
```

@@

#####**Graph 4.** Face Bounding Box Data: Box Plot 
```{r G4 Face Bounding Box Data, anchor="figure"}
ggplot(ALLmetaIMG, aes(x = factor(x = factor(type)), fill=type, y = size)) + geom_boxplot() + scale_y_continuous(trans='log') +xlab("Facial Recognition Softwares")+scale_x_discrete(limits=c("Google","Animetrics","Manual","Microsoft","Skybiometry"))+ylab("Size of Face Bounding Boxes")+guides(fill=FALSE)

@@

```{r I2, results=TRUE, fig.cap="2016_CT6_R02_JSousa_POR_vs_SGiraldo_COL_MS231_clip.0017.png", anchor="figure"}
imagesList2<-as.list("2016_CT6_R02_JSousa_POR_vs_SGiraldo_COL_MS231_clip.0017.png")
overlayGgplot(imagesList2, ALLmetaIMG, matchBox = FALSE)
```

@@

```{r I3, results=TRUE, fig.cap="2016_SC2_R01_ATomljanovic_AUS_vs_KBondarenko_UKR_WS1112_clip.0053.png", anchor="figure"}
imagesList3<-as.list("2016_SC2_R01_ATomljanovic_AUS_vs_KBondarenko_UKR_WS1112_clip.0053.png") 
overlayGgplot(imagesList3, ALLmetaIMG, matchBox = FALSE)
```

@@

```{r I4,results=TRUE, fig.cap="2016_RLA_R04_NDjokovic_SRB_vs_GSimon_FRA_MS401_clip.0003.png", anchor="figure"}
imagesList4<-as.list("2016_RLA_R04_NDjokovic_SRB_vs_GSimon_FRA_MS401_clip.0003.png")
overlayGgplot(imagesList4, ALLmetaIMG, matchBox = FALSE)
```

I4 shows the sensitivity of the software. We may consider this too sensitive as it recognised the eye and nose of the ballboy on the court.

##### Image 5.2016_SC2_R01_AKerber_GER_vs_MDoi_JPN_WS1332_clip.0083.png
```{r I5, results=TRUE, fig.cap="2016_SC2_R01_AKerber_GER_vs_MDoi_JPN_WS1332_clip.0083.png", anchor="figure"}
imagesList5<-("2016_SC2_R01_AKerber_GER_vs_MDoi_JPN_WS1332_clip.0083.png") 
overlayGgplot(imagesList5, ALLmetaIMG, matchBox = FALSE)
```

@@

```{r Type-Matches-Manual, anchor="table"}
library(descr, quietly = TRUE)
NoManual<-ALLmetaIMG %>% filter(ALLmetaIMG$type!="Manual")
NoManual<-NoManual %>% droplevels(NoManual$type)
Accuracy<-with(NoManual, table(NoManual$matchesManual, NoManual$type))
kable((Accuracy))
```

@@

```{r Image-Characteristics-Table, anchor="table"}
ImageCharacteristics <- ALLmetaIMG %>% filter(type=="Manual") %>% group_by(situation, bg, shotangle, detect) %>% dplyr::summarise(count=n())
ImageCharacteristics<-arrange(ImageCharacteristics, desc(count))
ImageCharacteristics<-ImageCharacteristics[1:10,]
kable(ImageCharacteristics, booktabs = TRUE)
```

`r figr("Image-Characteristics-Table", TRUE, type="Figure")`
 displays the feature combinations that produced the most potential face Bounding Boxes recognitions by all four softwares. The most common shot is a crowd shot.
The second row in the table with 830 faces recognised is more interesting than the first result. This useful scene is an image of a Court Player Close-Up in front of a Logo Wall, taken at Player Shoulder Height.

@@

```{r Typical-Best-Recognised, results=FALSE, anchor="figure"}
imagesList1<-as.list("2016_CT6_R01_JMillman_AUS_vs_DSchwartzman_ARG_MS159_clip.0013.png") 
overlayGgplot(imagesList1, ALLmetaIMG, matchBox = FALSE)
```

```{r Google-Errors-Characteristics-Table, anchor="Table"}
library(dplyr)

@@

GoogleNoMatch<-arrange(GoogleNoMatch, desc(count))
kable(GoogleNoMatch[1:10,], type = 'table')

@@

`r figr("Google-Errors-Characteristics-Table", TRUE, type="table")`
 shows that the potential Face Bounding Boxes Google returned were found in images that had the same characteristics of those that were manually annotated for faces.
<!-- This is shown by the same combinations of characteristics occuring in both Table 5 and Table 6.  keep as tables?-->
Without looking at each individual image this Table confirms that the potential Face Bounding Boxes Google located will likely be reasonable, and actually contain faces.

The best scenes for facial recongition have been found, given this information the following `r figr("Face-Characteristics-Table", TRUE, type="table")`
 considers the Characteristics of the individual faces found within those scenes. For this section we chose to consider only the faces that were manually annotated as players, with the intention of not basing results on recognitions of undesirable faces.

```{r Face-Characteristics-Table, anchor="Table"}
FaceCharacteristics <- ALLmetaIMGPlayers %>%   filter(type=="Google")%>% filter(matchesManual==TRUE)%>% group_by(visorhat, glasses, headangle, lighting, obscured, detect) %>% dplyr::summarise(count=n())
FaceCharacteristics<-arrange(FaceCharacteristics, desc(count))
FaceCharacteristics<-as.tbl(FaceCharacteristics[1:10,])
kable(FaceCharacteristics, type = 'table')
```

`r figr("Face-Characteristics-Table", TRUE, type="table")`
utilises a set of faces that were Manually annotated and also found by Google. 
Nine of the top ten facial characteristic combinations contained faces that were not wearing glasses.
The head angle describing the face angle was 'Other'^[Definition for headangle factor level 'Other' found in..]
for nine of the top ten facial characteristic combinations.

```{r Glasses, anchor="Figure"}
GlassesGoogle <- ALLmetaIMGPlayers %>% filter(type=="Google") %>%
filter(matchesManual==TRUE)

@@

```{r Glasses-Figure, anchor="Figure"}
ggplotProportion(ALLmetaIMGPlayers, glasses)
```
Graph 2 above shows that Google outperforms the other softwares, with or without glasses. When there are no glasses worn Google finds over 60% of the manually annotated faces.

```{r HeadAngles, anchor="Table"}
HeadangleGoogle <- ALLmetaIMGPlayers %>%   filter(type=="Google") %>%
filter(matchesManual==TRUE)

pander(table(HeadangleGoogle$headangle))
```
The amount of faces found by Google with the headangle "Other", is much larger than the amount of faces with the headangle "Profile" or "Front On". Therefore this should also be considered propotionately.

```{r Headangle-Figure, anchor="Figure"}
ggplotProportion(ALLmetaIMGPlayers, headangle)
```
Graph 3 diplays that Google performs much better in comparison to the other softwares when the headangle is Profile. This is outperforming unusually well, however this could also be due to the poor performance of the softwares in this circumstance.

@@

```{r Accuracy}
Accuracy<-with(ALLmetaIMG, table(ALLmetaIMG$matchesManual, ALLmetaIMG$type))
kable(Accuracy)
# make this a proportional table, would this be necessary if we have the UpSetR
```

@@

```{r Google-Characteristics-Table, anchor="table"}
GoogleCharacteristics <- ALLmetaIMG %>% filter(type=="Google") %>% filter(person=="Person") %>% filter(situation=="Court player close-up") %>% filter(bg=="Logo wall") %>% filter(detect=="Player") %>% group_by(person, situation, bg, shotangle, detect) %>% dplyr::summarise(count=n())
GoogleCharacteristics<-arrange(GoogleCharacteristics, desc(count))

kable(GoogleCharacteristics, format = "pandoc", booktabs = TRUE)
```

`r figr("Google-Characteristics-Table", TRUE, type="table")`
shows how many images with potential player's faces Google recognised. This diplays a vast gap between the amount of potential faces found given the different shot angles. Shoulder height is an optimum angle for a Face Bounding Box.

@@
